---
title: "Homework 4"
author: "Gurpreet Singh"
date: "April 15, 2018"
output:
  html_document: default
  word_document: default
---

```{r}
library(data.table)
library(mice)
library(knitr)
library(purrr)
library(ggplot2)
library(reshape)
library(DataExplorer)
library(statar)
library(corrplot)
library(psych)
library(caret)
library(pROC)
library(PerformanceAnalytics)
```



#### Data Exploration:
```{r}
train <- fread("https://raw.githubusercontent.com/gpsingh12/Data-621/master/Hw4/insurance_training_data.csv")
train <- train[,-c(1)]
var <- fread("https://raw.githubusercontent.com/gpsingh12/Data-621/master/Hw4/variables_ins_data.csv")
```


```{r}
str(train)
dim(train)
kable(head(train))

#psych::describe(train)

```

```{r}
table(train$PARENT1,useNA = 'ifany')
table(train$MSTATUS, useNA = 'ifany')
table(train$SEX, useNA = 'ifany')
table(train$EDUCATION, useNA = 'ifany')
table(train$JOB,useNA = 'ifany')
table(train$CAR_TYPE,useNA = 'ifany')
table(train$CAR_USE,useNA = 'ifany')
table(train$RED_CAR,useNA = 'ifany')
table(train$REVOKED,useNA = 'ifany')
table(train$URBANICITY,useNA = 'ifany')


train$INCOME <- as.numeric(gsub('[$,]', '', train$INCOME))
train$HOME_VAL <- as.numeric(gsub('[$,]', '', train$HOME_VAL))
train$BLUEBOOK <- as.numeric(gsub('[$,]', '', train$BLUEBOOK))
train$OLDCLAIM <- as.numeric(gsub('[$,]', '', train$OLDCLAIM))





miss <- data.frame(map_df(train, function(x) sum(is.na(x))))
miss <- miss[!(rowSums(miss) == 0), !(colSums(miss) == 0)]
miss

miss_plot <- plot_missing(train)
miss_plot

##https://stackoverflow.com/questions/8317231/elegant-way-to-report-missing-values-in-a-data-frame




train_out <- train[,-c(1:3,9,11:14,16,19,20,23,26)]
m1 <- melt(as.data.frame(train_out))

ggplot(m1,aes(x = variable,y = value)) + facet_wrap(~variable,scales = "free",ncol =4) + geom_boxplot()


p <-ggplot(m1, aes(x=variable, y=value)) + geom_boxplot(outlier.colour="red", outlier.shape=8,outlier.size=4)
p + coord_flip()
ggplot(m1, mapping = aes(x = value)) + geom_histogram(bins = 20) + facet_wrap(~variable, scales = 'free')

```










#### Data Preparation


Outlier Treatment


```{r}
#winsorize(train)

replace_outliers <- function(x) {
qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
caps <- quantile(x, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(x, na.rm = T)
x[x < (qnt[1] - H)] <- caps[1]
x[x > (qnt[2] + H)] <- caps[2]
 
return(x)
}

train$BLUEBOOK<-replace_outliers(train$BLUEBOOK)
train$OLDCLAIM<-replace_outliers(train$OLDCLAIM)
train$INCOME<-replace_outliers(train$INCOME)




train_out_2 <- train[,c(7,16,20,26,27,28)]
m2 <- melt(as.data.frame(train_out_2))
p <-ggplot(m2, aes(x=variable, y=value)) + geom_boxplot(outlier.colour="red", outlier.shape=8,outlier.size=4)
p + coord_flip()

```




Missing Values:

In data exploration section, we found five variabes with missing values.The percentage of missing values is 
5% - 6% of the original records. We will not remove these variables from analysis. Missing values will be treated for analysis rather than removing the variables from analysis.
Structure of variables age, yoj, income, home value and car age is numeric or integer rather than character. Using median imputation is a better practise in that case.

```{r}

train$AGE[is.na(train$AGE)] <- median(train$AGE, na.rm = TRUE)
train$YOJ[is.na(train$YOJ)] <- median(train$YOJ, na.rm = TRUE)
train$INCOME[is.na(train$INCOME)] <- median(train$INCOME, na.rm = TRUE)
train$HOME_VAL[is.na(train$HOME_VAL)] <- median(train$HOME_VAL, na.rm = TRUE)
train$CAR_AGE[is.na(train$CAR_AGE)] <- median(train$CAR_AGE, na.rm = TRUE)
```




New Variables
```{r}


train$PARENT1<- ifelse(train$PARENT1=="Yes",1,0)

train$MSTATUS<- ifelse(train$MSTATUS=="Yes",1,0)

train$SEX_m<- ifelse(train$SEX=="M",1,0)
train$SEX_f<- ifelse(train$SEX=="z_F",1,0)
train <- train[,-c("SEX")]

train$EDU_lesshs<- ifelse(train$EDUCATION=="<High School",1,0)
train$EDU_hs<- ifelse(train$EDUCATION=="z_High School",1,0)
train$EDU_bach<- ifelse(train$EDUCATION=="Bachelors",1,0)
train$EDU_ms<- ifelse(train$EDUCATION=="Masters",1,0)
train$EDU_phd<- ifelse(train$EDUCATION=="PhD",1,0)

train <- train[,-c("EDUCATION")]

train$JOB_dl<- ifelse(train$JOB=="Doctor"|train$JOB=="Lawyer",1,0)
train$JOB_prof<- ifelse(train$JOB=="Clerical"|train$JOB=="Manager"|train$JOB=="Professional",1,0)
train$JOB_bc<- ifelse(train$JOB=="z_Blue Collar",1,0)
train$JOB_st<- ifelse(train$JOB=="Student",1,0)
train$JOB_hm<- ifelse(train$JOB=="Home Maker",1,0)
train$JOB_nj<- ifelse(train$JOB==""|train$JOB == " ",1,0)

train <- train[,-c("JOB")]



train$CAR_USE_comm <- ifelse(train$CAR_USE=="Commercial",1,0)
train$CAR_USE_pvt<- ifelse(train$CAR_USE=="Private",1,0)
train <- train[,-c("CAR_USE")]

train$CAR_TYPE_suv <- ifelse(train$CAR_TYPE=="z_SUV",1,0)
train$CAR_TYPE_mvv <- ifelse(train$CAR_TYPE=="Minivan"|train$CAR_TYPE=="Van",1,0)
train$CAR_TYPE_sc <- ifelse(train$CAR_TYPE=="Sports Car",1,0)
train$CAR_TYPE_pt <- ifelse(train$CAR_TYPE=="Panel Truck",1,0)
train$CAR_TYPE_pu <- ifelse(train$CAR_TYPE=="Pickup",1,0)
train <- train[,-c("CAR_TYPE")]

train$RED_CAR<- ifelse(train$RED_CAR=="yes",1,0)




train$REVOKED<- ifelse(train$REVOKED=="Yes",1,0)


train$URBANICITY<- ifelse(train$URBANICITY=="Highly Urban/ Urban",1,0)



train <- train[,-c("SEX_f", "EDU_phd", "CAR_USE_pvt", "JOB_nj","CAR_TYPE_pu")]
```


Multicollinearity


Variables Income and home value shows collinearity. The correlation between the variables is not strong enough, we will retain the variables for analysis. For income variable we will create a new variable by transforming the variable using quantiles.

```{r}
library(corrplot)
x <- cor(train)
corrplot(x,  method="number")

cm <- data.frame(round(cor(train),2))
cm_high <- cm[cm]
```

Bucket Variables


#### Build Models

Prediction for TARGET_FLAG
Model1

Full model: original dataset with all the variables included in the dataset except newly generated transformed variables.

```{r}

train1<- train[,-c(2)]
model1 <- glm( TARGET_FLAG~.,family=binomial(link='logit'),data=train1)
summary(model1)



coefficients(model1)
predict1 <- predict(model1, type = 'response')
pred_1<- ifelse(predict1>0.5,1,0)
cm_model1 <- confusionMatrix(pred_1, train1$TARGET_FLAG, positive ="1")

par_m11 <- data.frame((cm_model1$byClass))
par_m12 <- data.frame((cm_model1$overall))
names(par_m11)<- c("parameters_model1")
names(par_m12)<- c("parameters_model1")
par_m1<- rbind(par_m11, par_m12)

kable(par_m1)

AIC(model1)

BIC(model1)



rc_m1 <- roc(train$TARGET_FLAG~ predict1)
auc(rc_m1)
plot(rc_m1)


```


Model 2
Backward Selection : Model 2 is an updated version of model 1 using backward elimination to remove insignificant variables.

```{r}
model2 <-  step(model1, direction="backward")
summary(model2)




coefficients(model2)
predict2 <- predict(model2, type = 'response')
pred_2<- ifelse(predict2>0.5,1,0)
cm_model2 <- confusionMatrix(pred_2, train1$TARGET_FLAG, positive ="1")

par_m21 <- data.frame((cm_model2$byClass))
par_m22 <- data.frame((cm_model2$overall))
names(par_m21)<- c("parameters_model2")
names(par_m22)<- c("parameters_model2")
par_m2<- rbind(par_m21, par_m22)

kable(par_m2)

AIC(model2)

BIC(model2)



rc_m2 <- roc(train$TARGET_FLAG~ predict2)
auc(rc_m2)
plot(rc_m2)


```



Model 3 

Forward Selection
```{r}

m_mdl<-glm(TARGET_FLAG ~ 1, data= train1)
full <- formula(lm(TARGET_FLAG~.,train1))

model3 <-step(m_mdl, direction='forward', scope=full)

summary(model3)



coefficients(model3)
predict3 <- predict(model3, type = 'response')
pred_3<- ifelse(predict3>0.5,1,0)
cm_model3 <- confusionMatrix(pred_3, train1$TARGET_FLAG, positive ="1")

par_m31 <- data.frame((cm_model3$byClass))
par_m32 <- data.frame((cm_model3$overall))
names(par_m31)<- c("parameters_model3")
names(par_m32)<- c("parameters_model3")
par_m3<- rbind(par_m31, par_m32)

kable(par_m3)

AIC(model3)

BIC(model3)



rc_m3 <- roc(train$TARGET_FLAG~ predict3)
auc(rc_m3)
plot(rc_m3)
```


### Model 4
```{r}
train2 <- train[,c("TARGET_FLAG", "KIDSDRIV","MVR_PTS", "REVOKED", "CAR_USE_comm", "TIF")]

model4 <- glm( TARGET_FLAG~.,family=binomial(link='logit'),data=train2)
summary(model4)



coefficients(model4)
predict4 <- predict(model4, type = 'response')
pred_4<- ifelse(predict4>0.5,1,0)
cm_model4 <- confusionMatrix(pred_4, train1$TARGET_FLAG, positive ="1")

par_m41 <- data.frame((cm_model4$byClass))
par_m42 <- data.frame((cm_model4$overall))
names(par_m41)<- c("parameters_model4")
names(par_m42)<- c("parameters_model4")
par_m4<- rbind(par_m41, par_m42)

kable(par_m4)

AIC(model4)
BIC(model4)

rc_m4 <- roc(train$TARGET_FLAG~ predict4)
auc(rc_m4)
plot(rc_m4)
```




```{r}
comp <- cbind(round(par_m1,3),round(par_m2,3),round(par_m3,3),round(par_m4,3))
comp


aic_df <- data.frame(AIC(model1), AIC(model2), AIC(model3),AIC(model4))
(aic_df)
bic_df <- data.frame(BIC(model1), BIC(model2), BIC(model3), BIC(model4))

(bic_df)
```





### Models for target amount

#### Model1

```{r}


#### Test Model:



test<- fread("https://raw.githubusercontent.com/gpsingh12/Data-621/master/Hw4/insurance-evaluation-data.csv")
dim(test)


table(test$PARENT1,useNA = 'ifany')
table(test$MSTATUS, useNA = 'ifany')
table(test$SEX, useNA = 'ifany')
table(test$EDUCATION, useNA = 'ifany')
table(test$JOB,useNA = 'ifany')
table(test$CAR_TYPE,useNA = 'ifany')
table(test$CAR_USE,useNA = 'ifany')
table(test$RED_CAR,useNA = 'ifany')
table(test$REVOKED,useNA = 'ifany')
table(test$URBANICITY,useNA = 'ifany')



test$INCOME <- as.numeric(gsub('[$,]', '', test$INCOME))
test$HOME_VAL <- as.numeric(gsub('[$,]', '', test$HOME_VAL))
test$BLUEBOOK <- as.numeric(gsub('[$,]', '', test$BLUEBOOK))
test$OLDCLAIM <- as.numeric(gsub('[$,]', '', test$OLDCLAIM))


miss <- data.frame(map_df(test, function(x) sum(is.na(x))))
miss <- miss[!(rowSums(miss) == 0), !(colSums(miss) == 0)]
miss



test$AGE[is.na(test$AGE)] <- median(test$AGE, na.rm = TRUE)
test$YOJ[is.na(test$YOJ)] <- median(test$YOJ, na.rm = TRUE)
test$INCOME[is.na(test$INCOME)] <- median(test$INCOME, na.rm = TRUE)
test$HOME_VAL[is.na(test$HOME_VAL)] <- median(test$HOME_VAL, na.rm = TRUE)
test$CAR_AGE[is.na(test$CAR_AGE)] <- median(test$CAR_AGE, na.rm = TRUE)




replace_outliers <- function(x) {
qnt <- quantile(x, probs=c(.25, .75), na.rm = T)
caps <- quantile(x, probs=c(.05, .95), na.rm = T)
H <- 1.5 * IQR(x, na.rm = T)
x[x < (qnt[1] - H)] <- caps[1]
x[x > (qnt[2] + H)] <- caps[2]
 
return(x)
}

test$BLUEBOOK<-replace_outliers(test$BLUEBOOK)
test$OLDCLAIM<-replace_outliers(test$OLDCLAIM)
test$INCOME<-replace_outliers(test$INCOME)



test$AGE[is.na(test$AGE)] <- median(test$AGE, na.rm = TRUE)
test$YOJ[is.na(test$YOJ)] <- median(test$YOJ, na.rm = TRUE)
test$INCOME[is.na(test$INCOME)] <- median(test$INCOME, na.rm = TRUE)
test$HOME_VAL[is.na(test$HOME_VAL)] <- median(test$HOME_VAL, na.rm = TRUE)
test$CAR_AGE[is.na(test$CAR_AGE)] <- median(test$CAR_AGE, na.rm = TRUE)





test$PARENT1<- ifelse(test$PARENT1=="Yes",1,0)

test$MSTATUS<- ifelse(test$MSTATUS=="Yes",1,0)

test$SEX_m<- ifelse(test$SEX=="M",1,0)
test$SEX_f<- ifelse(test$SEX=="z_F",1,0)
test <- test[,-c("SEX")]

test$EDU_lesshs<- ifelse(test$EDUCATION=="<High School",1,0)
test$EDU_hs<- ifelse(test$EDUCATION=="z_High School",1,0)
test$EDU_bach<- ifelse(test$EDUCATION=="Bachelors",1,0)
test$EDU_ms<- ifelse(test$EDUCATION=="Masters",1,0)
test$EDU_phd<- ifelse(test$EDUCATION=="PhD",1,0)

test <- test[,-c("EDUCATION")]

test$JOB_dl<- ifelse(test$JOB=="Doctor"|test$JOB=="Lawyer",1,0)
test$JOB_prof<- ifelse(test$JOB=="Clerical"|test$JOB=="Manager"|test$JOB=="Professional",1,0)
test$JOB_bc<- ifelse(test$JOB=="z_Blue Collar",1,0)
test$JOB_st<- ifelse(test$JOB=="Student",1,0)
test$JOB_hm<- ifelse(test$JOB=="Home Maker",1,0)
test$JOB_nj<- ifelse(test$JOB==""|test$JOB == " ",1,0)

test <- test[,-c("JOB")]



test$CAR_USE_comm <- ifelse(test$CAR_USE=="Commercial",1,0)
test$CAR_USE_pvt<- ifelse(test$CAR_USE=="Private",1,0)
test <- test[,-c("CAR_USE")]

test$CAR_TYPE_suv <- ifelse(test$CAR_TYPE=="z_SUV",1,0)
test$CAR_TYPE_mvv <- ifelse(test$CAR_TYPE=="Minivan"|test$CAR_TYPE=="Van",1,0)
test$CAR_TYPE_sc <- ifelse(test$CAR_TYPE=="Sports Car",1,0)
test$CAR_TYPE_pt <- ifelse(test$CAR_TYPE=="Panel Truck",1,0)
test$CAR_TYPE_pu <- ifelse(test$CAR_TYPE=="Pickup",1,0)
test <- test[,-c("CAR_TYPE")]

test$RED_CAR<- ifelse(test$RED_CAR=="yes",1,0)




test$REVOKED<- ifelse(test$REVOKED=="Yes",1,0)


test$URBANICITY<- ifelse(test$URBANICITY=="Highly Urban/ Urban",1,0)

test1<- test[,-c("SEX_f", "EDU_phd", "CAR_USE_pvt", "JOB_nj","CAR_TYPE_pu")]


pred_df<-predict(model2,type="response", newdata=test1)
pred_df <- ifelse(pred_df > 0.5, 1, 0)
```


```{r}
train_ta<- train[train$TARGET_FLAG==1,-c(1)]
model11 <- lm(TARGET_AMT~.,data=train_ta)
summary(model11)


library(leaps)
best.subset <- regsubsets(TARGET_AMT~., train_ta,nvmax=5)
best.subset.summary <- summary(best.subset)
best.subset.summary$outmat

best.subset.by.adjr2 <- which.max(best.subset.summary$adjr2)
best.subset.by.adjr2

##http://jadianes.me/best-subset-model-selection-with-R


train_ta1 <- train_ta[,c("TARGET_AMT" ,"BLUEBOOK","REVOKED", "MVR_PTS","CAR_AGE", "SEX_f")]
model11 <- lm(TARGET_AMT~.,data=train_ta1)
summary(model11)
AIC(model11)
arsq_m11 <-summary(model11)$adj.r.squared
BIC(model11)
logLik(model11)




model22<- lm(TARGET_AMT~.,data=train_ta)
model22 <-step(model22, direction="backward")
summary(model22)

arsq_m22 <-summary(model22)$adj.r.squared
AIC(model22)
BIC(model22)
logLik(model22)





m_mdl<-glm(TARGET_AMT ~ 1, data= train_ta)
full <- formula(lm(TARGET_AMT~.,train_ta))

model33 <-step(m_mdl, direction='forward', scope=full)

summary(model33)
arsq_m33 <-summary(model33)$adj.r.squared
AIC(model33)
BIC(model33)
logLik(model33)




aic_df <- data.frame(AIC(model11), AIC(model22), AIC(model33))
(aic_df)
bic_df <- data.frame(BIC(model11), BIC(model22), BIC(model33))

(bic_df)



```

